\documentclass[aps,prb,onecolumn,scriptaddress]{revtex4-2}
\usepackage{graphicx,amsmath,amsfonts,amssymb,bm,algorithm,algpseudocode,booktabs}
\usepackage[colorlinks=true, urlcolor=blue, linkcolor=blue, citecolor=blue]{hyperref}
\usepackage{tabularx}

\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\Id}{\mathbb{I}}
\newcommand{\eps}{\epsilon}
\newcommand{\Order}{\mathcal{O}}
\newcommand{\braket}[2]{\langle#1|#2\rangle}

\begin{document}
	
	\title{Numerical Implementation of MPS and DMRG}
	\author{Implementation Notes}
	\date{\today}
	\maketitle
	
	\section{Data Structures}
	
	An MPS with open boundary conditions is stored as a list of 3-index tensors:
	\begin{itemize}
		\item $A^{(i)}_{a_{i-1},\sigma_i,a_i}$: Left-canonical tensor at site $i$, dimensions $(D_{i-1} \times d \times D_i)$
		\item $B^{(i)}_{a_{i-1},\sigma_i,a_i}$: Right-canonical tensor, dimensions $(D_{i-1} \times d \times D_i)$  
		\item $M^{(i)}_{a_{i-1},\sigma_i,a_i}$: General (non-canonical) tensor during optimization
		\item Physical dimension $d$ (e.g., $d=2$ for spin-1/2)
	\end{itemize}
	
	\textit{Storage Convention}: For site $i$, tensor $T^{(i)}$ has shape $(D_{i-1}, d, D_i)$ where the middle index is physical.
	
	\textit{Matrix Reshape Convention (Critical)}: When flattening $T^{(i)}_{a_{i-1},\sigma_i,a_i}$ to matrix $\tilde{M}_{r,c}$ for QR/SVD, the row index $r$ must follow C-order (row-major) mapping:
	\begin{equation}
		r = a_{i-1} \cdot d + \sigma_i \quad \text{where } a_{i-1} \in [0,D_{i-1}-1], \sigma_i \in [0,d-1]
	\end{equation}
	Column index $c = a_i$ (right bond). This yields matrix shape $(D_{i-1}d) \times D_i$.
	
	\textit{Python/NumPy Implementation}:
	\begin{verbatim}
		# C-order (row-major) reshape - CRITICAL for index consistency
		M_tilde = M.reshape((D_left * d, D_right), order='C')
		# Verify: M_tilde[a_prev * d + sigma, a_curr] == M[a_prev, sigma, a_curr]
	\end{verbatim}
	\textit{Fortran/MATLAB}: Use \texttt{reshape(..., order='F')} with reversed index ordering $r = \sigma_i \cdot D_{i-1} + a_{i-1}$ (column-major), or transpose after C-order reshape.
	
	
	\section{Canonicalization Algorithms}
	
	\subsection{Left-Canonicalization (QR-based)}
	More stable than SVD for canonicalization when singular values aren't needed.
	
\begin{algorithm}[H]
	\caption{LeftCanonicalize (Numerically Stable)}
	\begin{algorithmic}[1]
		\Require MPS tensors $\{M^{(1)}, \dots, M^{(L)}\}$ with shapes $(D_{i-1}, d, D_i)$
		\Ensure Left-canonical $\{A^{(1)}, \dots, A^{(L)}\}$, last tensor $C$ (scalar or vector)
		\State $C \gets 1.0$ \Comment{Initial transfer matrix, $1 \times 1$}
		\For{$i = 1$ to $L-1$}
		\State \Comment{Contract incoming transfer matrix $C$ with current site}
		\State $\tilde{M}_{(a_{i-1}\sigma_i), a_i} \gets \sum_{a'_{i-1}} C_{a_{i-1}, a'_{i-1}} M^{(i)}_{a'_{i-1}, \sigma_i, a_i}$ 
		\State \Comment{For $i=1$: $C$ is $1\times 1$, so $\tilde{M}$ is simply $M^{(1)}$ reshaped to $(d) \times D_1$}
		\State \Comment{Matrix shape: $(D_{i-1}d) \times D_i$}
		\State \Comment{QR decomposition with column-pivoting for stability}
		\If{$D_{i-1} \cdot d < D_i$}
		\State \textbf{Warning:} Bond dimension $D_i$ exceeds physical capacity $D_{i-1}d$. 
		\State \phantom{\textbf{Warning:}} State is necessarily rank-deficient. Perform truncation or use random initialization.
		\EndIf
		\State $Q, R \gets \text{QR}(\tilde{M})$ \Comment{$Q$: $(D_{i-1}d) \times D_i$, $R$: $D_i \times D_i$}
		\State $A^{(i)}_{a_{i-1}, \sigma_i, a_i} \gets \text{reshape}(Q, (D_{i-1}, d, D_i))$
		\State $C \gets R$ \Comment{Transfer matrix to next site, shape $D_i \times D_i$}
		\EndFor
		\State \Comment{Final site: absorb last $C$}
		\State $A^{(L)}_{a_{L-1}, \sigma_L, 1} \gets \sum_{a'_{L-1}} C_{a_{L-1}, a'_{L-1}} M^{(L)}_{a'_{L-1}, \sigma_L, 1}$ \Comment{Shape: $(D_{L-1}, d, 1)$}
	\end{algorithmic}
\end{algorithm}

	
	\subsection{Mixed-Canonical Form at Bond $k$}
	Required for variational minimization.
	
	\begin{algorithm}[H]
		\caption{BringToBond\\
			\textit{Numerical Note:} RQ decomposition (used for right-canonicalization) is less stable than QR and may not be available in all libraries (e.g., PyTorch). Alternative: Reverse the site order, apply \textsc{LeftCanonicalize}, then reverse back.}
		\begin{algorithmic}[1]
			\Require MPS $\{M^{(i)}\}$, target bond $k$
			\Ensure Left-canonical $A^{(1..k)}$, right-canonical $B^{(k+1..L)}$, center matrix $S$
			\State Left-canonicalize sites $1 \dots k$ using QR sweeps
			\State Right-canonicalize sites $L \dots k+1$ using RQ sweeps (reverse direction)
			\State $S \gets$ matrix at bond $k$ (contains singular values if from SVD)
		\end{algorithmic}
	\end{algorithm}
	
	\section{Truncation and Compression}
	
	\subsection{SVD Compression with Tolerance}
	\begin{algorithm}[H]
		\caption{TruncateBond}
		\begin{algorithmic}[1]
			\Require Matrix $M$ (dimensions $m \times n$), max bond $D_{\max}$, tolerance $\eps$, \textbf{abs tolerance $\eps_{\text{abs}} \sim 10^{-14}$}
			\Ensure Truncated matrices $U, S, V^\dagger$ with $D \leq D_{\max}$
			\State $U, S, V^\dagger \gets \text{SVD}(M, \text{full\_matrices=False})$
			\State $s_i \gets \text{diag}(S)$ in descending order
			\State \Comment{Determine effective $D$ using relative and absolute tolerances}
			\State \Comment{Safety check: prevent $D=0$ when all singular values vanish}
			\If{$s_0 < \eps_{\text{abs}}$}
			\State \textbf{Warning:} All singular values below absolute tolerance.
			\State Set $D \gets 1$ (keep single largest, even if numerically zero).
			\Else
			\State $D \gets \min\left(D_{\max}, \text{count}\left(s_i > \max\left(\eps \cdot s_0, \eps_{\text{abs}}\right)\right)\right)$
			\State \textbf{where} $\eps_{\text{abs}} \sim 10^{-14}$ (double precision) prevents $D \to \infty$ when $s_0 \approx 0$
			\EndIf
			\State \Comment{Truncation error}
			\State $\eta \gets \sum_{i=D}^{r-1} s_i^2$ \Comment{Discarded weight}
			\State $U \gets U_{:,0:D-1}$, $S \gets S_{0:D-1,0:D-1}$, $V^\dagger \gets V^\dagger_{0:D-1,:}$
			\State \Return $U, S, V^\dagger, \eta$
		\end{algorithmic}
	\end{algorithm}
	
	\textit{Numerical Note}: For $\eps \sim 10^{-12}$, single-precision floats may underflow. Use double precision (complex128).
	
	\section{Ground State Algorithm: Single-Site DMRG}
	
	\subsection{Effective Hamiltonian Construction}
	The local Hamiltonian at site $i$ is built from pre-computed environments:
	
	\begin{align}
		L^{(i)}_{b_i}(a_{i-1}, a'_{i-1}) &= \sum_{a_{i-2}, a'_{i-2}} \sum_{\sigma, \sigma'} \sum_{b_{i-1}} A^{(i-1)*}_{a'_{i-2},\sigma',a'_{i-1}} \, L^{(i-1)}_{b_{i-1}}(a_{i-2}, a'_{i-2}) \, W^{(i-1)}_{b_{i-1},b_i}(\sigma, \sigma') \, A^{(i-1)}_{a_{i-2},\sigma,a_{i-1}} \\
		R^{(i)}_{b_i}(a_i, a'_i) &= \sum_{a_{i+1}, a'_{i+1}} \sum_{\sigma, \sigma'} \sum_{b_{i+1}} B^{(i+1)}_{a_i,\sigma,a_{i+1}} \, R^{(i+1)}_{b_{i+1}}(a_{i+1}, a'_{i+1}) \, W^{(i+1)}_{b_i,b_{i+1}}(\sigma, \sigma') \, B^{(i+1)*}_{a'_i,\sigma',a'_{i+1}}
	\end{align}	
	where $W^{(i)}$ is the MPO tensor at site $i$.
	
	\textit{Explicit Contraction Sequence (Implementation Template)}:
	For numerical implementation, express Eq.~(4) as explicit index contractions (zero-indexed):
	\begin{verbatim}
		# Left Environment Update (site i-1 -> i)
		# Input: L_old shape (D_W, D_{i-2}, D_{i-2}), A shape (D_{i-2}, d, D_{i-1})
		#        W shape (D_W, D_W, d, d) for bulk, or boundary-adjusted
		# Output: L_new shape (D_W, D_{i-1}, D_{i-1})
		
		# Step 1: Contract L_old with A and A*
		# Temp[a'_{i-2}, sigma, a_{i-1}, a'_{i-1}] = 
		#   sum_{a_{i-2}} L_old[b_{i-1}, a_{i-2}, a'_{i-2}] * A[a_{i-2}, sigma, a_{i-1}]
		# Step 2: Contract with W
		# Temp2[b_i, a_{i-1}, a'_{i-1}] = 
		#   sum_{b_{i-1}, sigma, sigma'} W[b_{i-1}, b_i, sigma, sigma'] * conj(A[a'_{i-2}, sigma', a'_{i-1}]) * Temp[...]
	\end{verbatim}
	Using \texttt{numpy.einsum} notation (Einstein summation):
	\begin{equation}
		\text{\texttt{L\_new = einsum('bxs,aso,bByr,arY->BYo', L\_old, A, W, conj(A))}}
	\end{equation}
	where indices map as: \texttt{b}=$b_{i-1}$, \texttt{x}=$a_{i-2}$, \texttt{s}=$a'_{i-2}$, \texttt{a}=$\sigma$, \texttt{r}=$a_{i-1}$, \texttt{B}=$b_i$, \texttt{Y}=$a'_{i-1}$, \texttt{y}=$\sigma'$.
	
	\textit{Verification}: After each update, check environment Hermiticity: $\|L - L^\dagger\|_{\max} < 10^{-14}$.
	
	
	\textit{Index Convention:} 
	\begin{itemize}
		\item $A^{(i)}_{a_{i-1},\sigma,a_i}$: MPS tensor, $a \in [0,D-1]$, $\sigma \in [0,d-1]$
		\item $W^{(i)}_{b_{i-1},b_i}(\sigma, \sigma') \equiv W^{(i)}_{b_{i-1},b_i,\sigma,\sigma'}$: MPO tensor, $b \in [0,D_W-1]$
		\item Environment $L^{(i)}_{b_i}$ is a $D_{i-1} \times D_{i-1}$ matrix for each MPO index $b_i$
	\end{itemize}
	
	\textbf{Index Range Convention:} 
	\begin{itemize}
		\item MPO bond indices $b_i \in \{0, 1, \dots, D_W-1\}$ (zero-indexed).
		\item For the Heisenberg MPO in Sec.~\ref{sec:mpo}: $b_{\text{left}} = 4$, $b_{\text{right}} = 0$.
		\item Verify array bounds: $W^{(i)}$ has shape $(D_W, D_W, d, d)$ for bulk sites.
	\end{itemize}
	
	\textit{Index Dependency Verification}:
	\begin{itemize}
		\item $L^{(i)}$ depends on sites $1 \dots i-1$, hence uses $W^{(i-1)}$ (not $W^{(i)}$)
		\item $R^{(i)}$ depends on sites $i+1 \dots L$, hence uses $W^{(i+1)}$ (not $W^{(i)}$)  
		\item \textbf{Common bug}: Off-by-one error in MPO index. Verify: $L^{(2)}$ uses $W^{(1)}$, $R^{(L-1)}$ uses $W^{(L)}$.
	\end{itemize}
	\textit{Boundary Consistency}: For open boundaries, the MPO starts and ends with identity operators. Ensure $W^{(1)}$ has shape $(1, D_W, d, d)$ and $W^{(L)}$ has shape $(D_W, 1, d, d)$, or handle boundary vectors explicitly as $1 \times D_W$ and $D_W \times 1$ matrices.
	
	\subsection{Local Optimization}
	Solve the eigenvalue problem for the center tensor $M^{(i)}_{\alpha_{i-1},\alpha_i}$ reshaped as vector $v$ of length $d D_{i-1} D_i$:
	
	\begin{equation}
		\mathcal{H}_{\text{eff}} \, v = E v
	\end{equation}
	
	where 
	\begin{equation}
		\mathcal{H}_{\text{eff}} = \sum_{b_{i-1}, b_i} L_{b_{i-1}} \otimes W^{(i)}_{b_{i-1},b_i} \otimes R_{b_i}
	\end{equation}
	
	\textit{Implementation (Matrix-Free)}: Use Lanczos or Davidson with the following tensor network contractions (e.g., using \texttt{numpy.einsum}):
	
	\textit{Lanczos stopping criterion}: Stop when eigenvalue residual $\|\mathcal{H}_{\text{eff}}v - Ev\| < \delta$ where $\delta \sim 10^{-10}$ or when $N_{\text{iter}} > 20-50$.
		
	\textbf{Input:} Vector $v$ of length $d D_{i-1} D_i$ (reshaped as $V_{a_{i-1},\sigma_i,a_i}$)\\
	
	\textbf{Output:} $(\mathcal{H}_{\text{eff}}v)_{a_{i-1},\sigma_i,a_i}$
	
	\begin{algorithm}[H]
		\caption{ApplyEffectiveHamiltonian (Matrix-Free)}
		\begin{algorithmic}[1]
			\Require Vector $v$ (flattened from $V_{a_{i-1},\sigma_i,a_i}$), $L$ shape $(D_W, D_{i-1}, D_{i-1})$, $R$ shape $(D_W, D_i, D_i)$, $W$ shape $(D_W, D_W, d, d)$
			\Ensure $(\mathcal{H}_{\text{eff}}v)_{a_{i-1},\sigma'_i,a_i}$
			\State Reshape $v \to V_{a_{i-1},\sigma_i,a_i}$ \Comment{Shape $(D_{i-1}, d, D_i)$}
			\State \textbf{Step 1 (Left)}: $X_{b_{i-1},a_{i-1},\sigma_i,a_i} = \sum_{a'_{i-1}} L_{b_{i-1},a_{i-1},a'_{i-1}} V_{a'_{i-1},\sigma_i,a_i}$
			\State \phantom{\textbf{Step 1}} \textit{einsum}: \texttt{'bxy,ysz->bxsz'}
			\State \textbf{Step 2 (MPO)}: $Y_{b_i,a_{i-1},\sigma'_i,a_i} = \sum_{b_{i-1},\sigma_i} W_{b_{i-1},b_i,\sigma_i,\sigma'_i} X_{b_{i-1},a_{i-1},\sigma_i,a_i}$  
			\State \phantom{\textbf{Step 2}} \textit{einsum}: \texttt{'bBst,bxsz->Bxsz'} (swap to \texttt{'bBts'} if index order differs)
			\State \textbf{Step 3 (Right)}: $(\mathcal{H}v)_{a_{i-1},\sigma'_i,a_i} = \sum_{b_i, a'_i} R_{b_i,a_i,a'_i} Y_{b_i,a_{i-1},\sigma'_i,a'_i}$
			\State \phantom{\textbf{Step 3}} \textit{einsum}: \texttt{'BYz,Bxsz->xsY'} 
			\State \Return Flatten result to vector of length $d D_{i-1} D_i$
		\end{algorithmic}
	\end{algorithm}
	\textit{Cost}: $\Order(D_W D_{i-1} D_i d^2 + D_W^2 D_{i-1} D_i d)$. For sparse MPOs (Heisenberg), exploit zero blocks to reduce $D_W^2 \to D_W^{\text{eff}}$.
	
		
	\subsection{Environment Initialization}
	Before sweeping, initialize boundary environments consistent with the MPO structure:
	\begin{align}
		L^{(1)}_{b_0}(a_0, a'_0) &= \delta_{b_0, b_{\text{left}}} \delta_{a_0, a'_0} & \text{(Left boundary: $1 \times 1$ matrices)} \\
		R^{(L)}_{b_L}(a_L, a'_L) &= \delta_{b_L, b_{\text{right}}} \delta_{a_L, a'_L} & \text{(Right boundary: $1 \times 1$ matrices)}
	\end{align}
	where $b_{\text{left}}$ and $b_{\text{right}}$ are the MPO indices selecting the Hamiltonian scalar (e.g., for the Heisenberg MPO in Sec.~\ref{sec:mpo}, $b_{\text{left}}=4$ (last row), $b_{\text{right}}=0$ (first column)).
	
	\subsection{Complex Arithmetic Handling}
	Even for real Hamiltonians (e.g., Heisenberg), maintain \texttt{complex128} dtype throughout:
	\begin{itemize}
		\item \textbf{Left environment:} Uses complex conjugate only (no transpose):
		$L^{(i)} \gets A^{(i-1)*} \cdots$ (Python: \texttt{A.conj()}, \textbf{not} \texttt{A.T.conj()}).
		\item \textbf{Right environment:} Uses conjugate transpose: $B^{(i+1)\dagger} = (B^{(i+1)*})^T$ 
		(Python: \texttt{B.conj().T} or \texttt{B.T.conj()}).
		\item \textbf{Verification:} The density matrix $\rho = M M^\dagger$ must be Hermitian. 
		Check $\|\rho - \rho^\dagger\|_{\max} < 10^{-14}$ to detect index ordering bugs.
	\end{itemize}
	
	\subsection{Datatype and Library Requirements}
	
	\textit{Floating Point Precision}: All tensors must use \texttt{complex128} (double precision) even for real Hamiltonians. Single precision (\texttt{float32}) leads to underflow in $\epsilon \sim 10^{-12}$ truncations.
	
	\textit{QR Decomposition}: Standard \texttt{numpy.linalg.qr} lacks column-pivoting. Use:
	\begin{verbatim}
		from scipy.linalg import qr
		Q, R, P = qr(M, pivoting=True, mode='economic')
		# Reorder Q according to permutation P if necessary
	\end{verbatim}
	Ensure diagonal elements of $R$ have non-negative real parts (multiply column $j$ of $Q$ by $\text{sign}(R_{jj})$ if needed).
	
	\textit{SVD Truncation}: Explicitly sort singular values descending:
	\begin{verbatim}
		U, s, Vh = svd(M, full_matrices=False)
		# Ensure descending (NumPy default) - reverse if using different library
		idx = np.argsort(s)[::-1]
		s, U, Vh = s[idx], U[:, idx], Vh[idx, :]
	\end{verbatim}
	
	\subsection{Sweep Algorithm}
	
\begin{algorithm}[H]
	\caption{SingleSiteDMRG (with Gauge Control)}
	\begin{algorithmic}[1]
		\Require Hamiltonian MPO $\{W^{(i)}\}$, initial MPS guess $\{M^{(i)}\}$, max bond $D_{\max}$, tolerance $\eps$
		\Ensure Ground state energy $E_0$ and optimized MPS in mixed-canonical form
		\State Right-canonicalize initial MPS using RQ sweeps (or reverse QR)
		\State Initialize all $R^{(i)}$ environments via backward contraction from site $L$
		\State $L^{(1)} \gets$ left boundary tensor (see initialization above)
		\While{not converged}
		\State \textbf{Right sweep} ($i=1$ to $L-1$):
		\For{$i=1$ to $L-1$}
		\State Build $\mathcal{H}_{\text{eff}}$ from $L^{(i)}$, $R^{(i)}$, $W^{(i)}$
		\State Solve $\mathcal{H}_{\text{eff}} v = E v$ using matrix-free Lanczos/Davidson
		\State \Comment{Use the tensor contraction sequence in Sec.~4.2; never construct $\mathcal{H}_{\text{eff}}$ explicitly}
		\State \Comment{Dimensions: $d D_{i-1} D_i$ can be $\sim 10^4$-$10^6$, matrix size $\sim$ TB if explicit}
		\State Reshape $v \to M^{(i)}_{a_{i-1},\sigma_i,a_i}$
		\State \Comment{Gauge update: SVD and shift center to the right}
		\State Reshape $M^{(i)} \to \tilde{M}_{(a_{i-1}\sigma_i), a_i}$ (matrix $(dD_{i-1}) \times D_i$)
		\State $U, S, V^\dagger \gets \text{SVD}(\tilde{M})$, truncate to $D_{\max}$
		\State $D_{\text{new}} \gets \min(D_{\max}, \text{len}(S))$ \Comment{Actual dimension after truncation}
		\State $A^{(i)} \gets \text{reshape}(U, (D_{i-1}, d, D_{\text{new}}))$ \Comment{Left-canonical}
		\State \Comment{Absorb singular values into next site (center moves right)}
		\State $M^{(i+1)}_{a_i,\sigma_{i+1},a_{i+2}} \gets \sum_{a'_i, a''_i} \text{diag}(S)_{a_i,a'_i} V^\dagger_{a'_i,a''_i} M^{(i+1)}_{a''_i,\sigma_{i+1},a_{i+2}}$
		\State Update $L^{(i+1)}$ using $A^{(i)}$ and $W^{(i)}$ (see Section 4.1, left environment update)
		\EndFor
		\State \textbf{Left sweep} ($i=L$ to $2$):
		\For{$i=L$ to $2$}
		\State Solve eigenproblem using updated environments
		\State Reshape $M^{(i)} \to \tilde{M}_{a_{i-1}, (\sigma_i a_i)}$ (matrix $D_{i-1} \times (dD_i)$)
		\State $U, S, V^\dagger \gets \text{SVD}(\tilde{M})$, truncate to $D_{\max}$
		\State $D_{\text{new}} \gets \min(D_{\max}, \text{len}(S))$ \Comment{Actual dimension after truncation}
		\State $B^{(i)} \gets \text{reshape}(V^\dagger, (D_{\text{new}}, d, D_i))$ \Comment{Right-canonical}
		\State \Comment{Absorb singular values into previous site (center moves left)}
		\State $M^{(i-1)}_{a_{i-2},\sigma_{i-1},a_{i-1}} \gets \sum_{a'_{i-1}, a''_{i-1}} M^{(i-1)}_{a_{i-2},\sigma_{i-1},a''_{i-1}} U_{a''_{i-1},a'_{i-1}} \text{diag}(S)_{a'_{i-1},a_{i-1}}$
		\State Update $R^{(i-1)}$ using $B^{(i)}$ and $W^{(i)}$ (see Section 4.1, right environment update)
		\EndFor
		\State Check convergence (ALL must hold):
		\State \quad 1. Energy: $|\Delta E| = |E_n - E_{n-1}| < \eps_E$ (e.g., $10^{-8}$ per site)
		\State \quad 2. \textbf{Variance}: $\sigma^2 = \bra{\psi}H^2\ket{\psi} - E^2 < \eps_{\sigma}$ (stricter, $\sim 10^{-6}$)
		\State \quad 3. \textbf{State convergence}: $\max_i \|A^{(i)}_{\text{new}} - A^{(i)}_{\text{old}}\|_{\text{Frob}} < \eps_{\text{tensor}}$
		\State \quad 4. \textbf{Truncation stability}: $\eta_{\text{max}} < 10^{-7}$ (discarded weight across all bonds)
		\EndWhile
	\end{algorithmic}
\end{algorithm}

	\textit{Critical Implementation Note — The Local Minimum Problem}:
	Single-site updates with fixed bond dimension $D$ cannot escape local minima during phase transitions or for systems with symmetries. Two mitigation strategies:
	\begin{enumerate}
		\item \textbf{Two-Site DMRG}: Use Algorithm 5 for the first 3-5 sweeps to dynamically increase $D$, then switch to single-site for speed.
		\item \textbf{Perturbation/Noise Injection}: Before truncation (line 12), add random noise to the singular values or density matrix:
		\begin{equation}
			\rho \to \rho + \alpha \cdot \mathbb{I}, \quad \alpha \sim 10^{-5} \text{ to } 10^{-3}
		\end{equation}
		then re-diagonalize. Decrease $\alpha$ linearly with sweep number.
	\end{enumerate}
	Without these, single-site DMRG may converge to excited states with incorrect energies.
	
	\section{Two-Site DMRG (Robust against local minima)}
	
	For systems with symmetries or near phase transitions, single-site DMRG may get stuck. Two-site updates allow bond dimension adaptation.
	
	\begin{algorithm}[H]
		\caption{TwoSiteUpdate}
		\begin{algorithmic}[1]
			\Statex \textit{Effective Hamiltonian on two-site space:}
			\Statex $\mathcal{H}_{\text{eff}}^{(i,i+1)} = \sum_{b_{i-1}, b_{i+1}} L_{b_{i-1}} \otimes W^{(i)}_{b_{i-1},b_i} \otimes W^{(i+1)}_{b_i,b_{i+1}} \otimes R_{b_{i+1}}$
			\Statex where the sum contracts over the intermediate MPO bond $b_i \in [0, D_W-1]$.
			\Statex Dimensions: $d^2 D_{i-1} D_{i+1} \times d^2 D_{i-1} D_{i+1}$.
			\Require Current bond dimensions $D_{i-1}, D_i$, max bond $D_{\max}$, tolerance $\eps$					
			\State Contract: $\Theta_{\alpha_{i-1},\sigma_i,\sigma_{i+1},\alpha_{i+1}} = \sum_{\alpha_i} A^{(i)}_{\alpha_{i-1},\sigma_i,\alpha_i} B^{(i+1)}_{\alpha_i,\sigma_{i+1},\alpha_{i+1}}$
			\State Reshape: $\Theta \to \Theta_{(\alpha_{i-1}\sigma_i), (\sigma_{i+1}\alpha_{i+1})}$ (matrix of size $(d D_{i-1}) \times (d D_{i+1})$)
			\State Build $\mathcal{H}_{\text{eff}}$ on combined two-site space (dimensions $d^2 D_{i-1} D_{i+1}$)
			\State \Comment{Use matrix-free approach: apply $\mathcal{H}_{\text{eff}}$ via tensor contractions as in Alg.~4}
			\State \Comment{Never construct full matrix: dimension $d^2 D_{i-1} D_{i+1}$ can exceed $10^4$-$10^6$}
			\State Solve for lowest eigenvector $\tilde{\Theta}$ (size $d^2 D_{i-1} D_{i+1}$) using Lanczos/Davidson
			\State Reshape $\tilde{\Theta} \to$ matrix $(d D_{i-1}) \times (d D_{i+1})$
			\State SVD: $\tilde{\Theta} = U S V^\dagger$
			\State Truncate to $D \leq D_{\max}$ singular values using \textsc{TruncateBond}
			\State $A^{(i)} \gets \text{reshape}(U, (D_{i-1}, d, D))$ \Comment{Left-canonical}
			\State $B^{(i+1)} \gets \text{reshape}(S V^\dagger, (D, d, D_{i+1}))$ \Comment{Right-canonical}
			\State \Return Updated tensors, truncation error $\eta = \sum_{\text{discarded}} s_i^2$
		\end{algorithmic}
	\end{algorithm}
	
	\textit{Computational Cost}: Two-site update scales as $\Order(D^3 d^3)$ vs $\Order(D^3 d)$ for single-site, but allows $D$ to grow temporarily to $d D$ before truncation.
	
	\section{MPO Construction for Hamiltonians}
	\label{sec:mpo}
	
	
	\subsection{Nearest-Neighbor Heisenberg Model}
	MPO bond dimension $D_W = 5$. Define operator-valued matrices $\hat{W}^{(i)}$:
	
	\begin{equation}
		\hat{W}^{(i)} = \begin{pmatrix}
			\hat{\Id} & \cdot & \cdot & \cdot & \cdot \\
			\hat{S}^+ & \cdot & \cdot & \cdot & \cdot \\
			\hat{S}^- & \cdot & \cdot & \cdot & \cdot \\
			\hat{S}^z & \cdot & \cdot & \cdot & \cdot \\
			-h\hat{S}^z & \frac{J}{2}\hat{S}^- & \frac{J}{2}\hat{S}^+ & J_z \hat{S}^z & \hat{\Id}
		\end{pmatrix}
	\end{equation}
	
Boundary conditions (0-indexed array convention):
\begin{itemize}
	\item Site 1 (left boundary): $\hat{W}^{(1)}$ is a $1 \times D_W$ row vector selecting the last index (index 4 for $D_W=5$):
	\begin{equation}
		\hat{W}^{(1)}_{0,b} = \begin{cases} \text{row } b=4 \text{ of bulk matrix} & \text{if } b=4 \\ 0 & \text{otherwise} \end{cases}
	\end{equation}
	Shape: $(1, D_W, d, d)$. \textbf{Code}: \texttt{W[0, 4, :, :] = bulk\_W[4, 4, :, :]} (Hamiltonian scalar component).
	
	\item Site $L$ (right boundary): $\hat{W}^{(L)}$ is a $D_W \times 1$ column vector selecting the first index (index 0):
	\begin{equation}
		\hat{W}^{(L)}_{b,0} = \begin{cases} \text{column } b=0 \text{ of bulk matrix} & \text{if } b=0 \\ 0 & \text{otherwise} \end{cases}
	\end{equation}
	Shape: $(D_W, 1, d, d)$. \textbf{Code}: \texttt{W[0, 0, :, :] = bulk\_W[0, 0, :, :]} (Identity component).
\end{itemize}

\textit{Index Verification}: MPO bond indices $b_i \in \{0, 1, \dots, D_W-1\}$. For Heisenberg with $D_W=5$:
\begin{itemize}
	\item $b_{\text{left}} = 4$ (index of $\hat{\Id}$ in last row)
	\item $b_{\text{right}} = 0$ (index of $\hat{\Id}$ in first column)
\end{itemize}
Boundary environments: $L^{(1)}_{b_0} = \delta_{b_0, 4}$, $R^{(L)}_{b_L} = \delta_{b_L, 0}$.

	
	\subsection{Compression of MPO-MPS Product}
	When applying MPO to MPS, bond dimension grows to $D \cdot D_W$. Compress variationally:
	
\begin{algorithm}[H]
	\caption{VariationalMPSCompression}
	\begin{algorithmic}[1]
		\Require Target MPS $\ket{\psi}$ (bond $D_{\psi}$), initial guess $\ket{\phi}$ (bond $D_{\phi} < D_{\psi}$)
		\Ensure $\ket{\phi} \approx \ket{\psi}$ with minimal $\|\phi - \psi\|^2$
		\State Bring $\ket{\phi}$ to mixed-canonical form at bond $k=1$ (right-canonicalized)
		\State Compute initial environments $L_{\psi\phi}, R_{\psi\phi}$ (overlaps between $\psi$ and $\phi$)
		\While{not converged}
		\For{$i=1$ to $L-1$} \Comment{Right sweep}
		\State \Comment{Solve local least-squares: minimize $\|\phi\|^2 - 2\text{Re}\braket{\phi}{\psi}$}
		\State $N_{a_{i-1},a_i} \gets \sum_{b_i} L^{(i)}_{b_i} R^{(i)}_{b_i}$ \Comment{Metric, shape $D_{\phi} \times D_{\phi}$}
		\State $T_{a_{i-1},\sigma_i,a_i} \gets \sum_{\{a'\}} L_{\psi\phi} \cdot M^{(i)}_{\psi} \cdot R_{\psi\phi}$ \Comment{Overlap term (see ITensor docs)}
		\State $M^{(i)}_{\phi} \gets \text{Reshape}(N^{-1} \cdot \text{Flatten}(T))$ \Comment{Pseudo-inverse if singular}
		\State SVD and shift gauge right (as in DMRG)
		\EndFor
		\State Left sweep (analogous)
		\EndWhile
	\end{algorithmic}
\end{algorithm}
	
	\section{Model Examples}
	
	\subsection{Spin-1/2 Heisenberg Chain}
	The antiferromagnetic Heisenberg model is the paradigmatic test case for DMRG algorithms. For an open chain of length $L$, the Hamiltonian is
	\begin{equation}
		H = J \sum_{i=1}^{L-1} \vec{S}_i \cdot \vec{S}_{i+1} = J \sum_{i=1}^{L-1} \left[ \frac{1}{2}\left(S_i^+ S_{i+1}^- + S_i^- S_{i+1}^+\right) + S_i^z S_{i+1}^z \right],
	\end{equation}
	with $J>0$. The ground state exhibits quasi-long-range antiferromagnetic order (power-law decay of correlations $\sim (-1)^r/r$ at criticality). 
	
	\textbf{MPO Representation:} As detailed in Sec.~\ref{sec:mpo}, this model admits a compact MPO with bond dimension $D_W=5$. The operator-valued matrices are
	\begin{equation}
		\hat{W}^{(i)} = \begin{pmatrix}
			\hat{\Id} & \cdot & \cdot & \cdot & \cdot \\
			\hat{S}^+ & \cdot & \cdot & \cdot & \cdot \\
			\hat{S}^- & \cdot & \cdot & \cdot & \cdot \\
			\hat{S}^z & \cdot & \cdot & \cdot & \cdot \\
			\cdot & \frac{J}{2}\hat{S}^- & \frac{J}{2}\hat{S}^+ & J\hat{S}^z & \hat{\Id}
		\end{pmatrix},
	\end{equation}
	with boundary vectors selecting the bottom-left and top-right elements.
	
	\textbf{Numerical Notes:} The Heisenberg ground state is not exactly representable by a finite-$D$ MPS. Due to its critical nature, the entanglement entropy scales logarithmically with system size, requiring bond dimensions $D \sim L^{\alpha}$ (with $\alpha \approx 1/6$) to maintain accuracy. Results must be extrapolated in the truncation error $\eta$ or $1/D$.
	
	\subsection{AKLT Model (Exact MPS Representation)}
	The Affleck-Kennedy-Lieb-Tasaki (AKLT) model is an $S=1$ spin chain with Hamiltonian
	\begin{equation}
		H_{\text{AKLT}} = \sum_{i=1}^{L-1} \left[ \vec{S}_i \cdot \vec{S}_{i+1} + \frac{1}{3} (\vec{S}_i \cdot \vec{S}_{i+1})^2 \right].
	\end{equation}
	Unlike the Heisenberg model, the AKLT ground state is gapped and can be represented \textit{exactly} as a translationally invariant MPS with bond dimension $D=2$.
	
	\textbf{Physical Picture:} Each $S=1$ site is constructed from two $S=1/2$ spins (auxiliary space) in a symmetric triplet state. Neighboring sites are linked by singlet bonds between auxiliary spins, followed by a projection onto the physical $S=1$ subspace.
	
	\textbf{Explicit MPS Matrices:} The left-canonical matrices $A^{\sigma}$ (where $\sigma \in \{+,0,-\}$ corresponds to $m_z = +1, 0, -1$) are given by
	\begin{align}
		A^{+} &= \begin{pmatrix} 0 & \sqrt{2/3} \\ 0 & 0 \end{pmatrix}, \\
		A^{0} &= \begin{pmatrix} -1/\sqrt{3} & 0 \\ 0 & 1/\sqrt{3} \end{pmatrix}, \\
		A^{-} &= \begin{pmatrix} 0 & 0 \\ -\sqrt{2/3} & 0 \end{pmatrix}.
	\end{align}
	These satisfy the left-normalization condition $\sum_{\sigma} A^{\sigma\dagger} A^{\sigma} = \mathbb{I}_2$.
	
	\textit{Verification Test}: After implementing the MPS class, verify:
	\begin{equation}
		\sum_{\sigma \in \{+,0,-\}} A^{\sigma\dagger} A^{\sigma} = \mathbb{I}_2, \quad 
		\sum_{\sigma} A^{\sigma} A^{\sigma\dagger} = \frac{2}{3}\mathbb{I}_2 + \frac{1}{3}\sigma_z \neq \mathbb{I}
	\end{equation}
	The left-canonical condition is satisfied; the right-canonical is not (this is an \textit{injective} MPS). For two-site DMRG initialization with $D=2$, the energy should converge to exactly $E_0 = -(L-1)\cdot \frac{2}{3}$ with truncation error $\eta \equiv 0$.
	
	\textbf{Properties:}
	\begin{itemize}
		\item \textbf{Correlation Functions:} Spin-spin correlations decay exponentially as $\langle S^z_i S^z_j \rangle \sim (-1/3)^{|i-j|}$ with correlation length $\xi = 1/\ln 3 \approx 0.91$.
		\item \textbf{Hidden Order:} The model exhibits a non-local "string order" $\langle S^z_i \exp(i\pi \sum_{k=i+1}^{j-1} S^z_k) S^z_j \rangle = -4/9$ for $|i-j|>2$, indicating topological order.
		\item \textbf{Open Boundaries:} For OBC, the exact ground state is $\ket{\psi} = \sum_{\{\sigma\}} A^{\sigma_1} \cdots A^{\sigma_L} \ket{\sigma_1 \cdots \sigma_L}$, resulting in two free $S=1/2$ edge spins (degenerate ground state manifold).
	\end{itemize}
	
	This model serves as an excellent benchmark for DMRG implementations: with $D=2$, the algorithm should reproduce the exact energy $E_0 = -(L-1) \cdot 2/3$ (for $J=1$) and exponentially decaying correlations without truncation error.
	
	\section{Convergence Diagnostics}
	
	\begin{itemize}
		\item \textbf{Energy convergence}: $\Delta E = |E_{n} - E_{n-1}| < 10^{-8}$ (per site)
		\item \textbf{Truncation error}: Monitor $\eta = \sum_{\text{discarded}} s_i^2$. For accurate results, extrapolate $E(\eta) \to \eta \to 0$.
		\item \textbf{Variance}: $\sigma^2 = \bra{\psi}H^2\ket{\psi} - \bra{\psi}H\ket{\psi}^2$. Should be $\sim D^{-2}$ for gapped systems.
		\item \textbf{Entanglement entropy}: Check $S = -\sum s_i^2 \ln s_i^2$ saturation across bonds.
	\end{itemize}
	
\textit{Computational Note on Variance:} Computing $\langle H^2 \rangle$ by squaring the MPO 
(bond dimension $D_W^2$) is inefficient. Instead:
\begin{enumerate}
	\item Compute $\ket{\phi} = H\ket{\psi}$ via MPO-MPS multiplication (cost $\sim LD_WD^3$).
	\item Compute $\langle H^2 \rangle = \braket{\phi}{\phi}$ (MPS overlap).
	\item Reuse $\langle\psi|H|\psi\rangle$ from DMRG energy for $E^2$ term.
\end{enumerate}
This avoids the $D_W^2$ scaling of explicit $H^2$ construction.

		
	\section{Implementation Verification Checklist}
	\label{sec:verification}
	
	The following phased testing protocol validates correct implementation of each subsystem before proceeding to full DMRG. Implement in order; failures at early stages indicate index ordering or conjugation bugs that must be fixed before testing variational optimization.
	
	\begin{table}[htbp]
		\centering
		\caption{Phased Implementation Tests}
		\label{tab:verification}
		\begin{tabularx}{\textwidth}{@{}l>{\raggedright\arraybackslash}p{3.5cm}X@{}}
			\toprule
			\textbf{Phase} & \textbf{Test} & \textbf{Expected Result} \\
			\midrule
			1. \texttt{Canonicalization} & 
			QR on random MPS $\to$ \textsc{LeftCanonicalize} & 
			Check $\sum_\sigma A^{(i)\sigma\dagger} A^{(i)\sigma} = \mathbb{I}_{D_i}$ for $i<L$ (within $10^{-14}$). Last site is vector $(D_{L-1}d) \times 1$. \\
			\midrule
			2. \texttt{Truncation} & 
			SVD compression of AKLT state ($D_{\max}=2$) & 
			$\eta = 0$ (exact representation maintained). Bond dimension remains $D=2$. \\
			\midrule
			3. \texttt{Environments} & 
			Build $L^{(i)}, R^{(i)}$ for AKLT with identity MPO & 
			Environments proportional to identity on auxiliary space. Verify $L^{(i)} \propto \delta_{a_{i-1},a'_{i-1}}$. \\
			\midrule
			4. \texttt{SingleSite} & 
			AKLT model, $L=10$, $D=2$, 2 sweeps & 
			Ground state energy $E_0 = -(L-1) \cdot 2/3 = -6.0$ (machine precision, $\Delta E < 10^{-12}$). \\
			\midrule
			5. \texttt{TwoSite} & 
			Heisenberg $L=20$, $D: 4\to20$, $\eps=10^{-10}$ & 
			Energy extrapolates to $-8.772$ (Bethe ansatz). Bond dimension adapts dynamically. \\
			\midrule
			6. \texttt{Convergence} & 
			Run single-site DMRG on gapped system & 
			Variance $\sigma^2 < 10^{-6}$ (strictest test). Check $\sigma^2 \sim D^{-2}$ scaling. \\
			\bottomrule
		\end{tabularx}
	\end{table}

	
	\subsection{Common Failure Modes}
	
	If tests fail, check these typical implementation errors:
	
	\begin{itemize}
		\item \textbf{Phase 1 fails (non-orthogonal $A$)}: Reshape index ordering incorrect. Verify row-major vs column-major convention (Section 1).
		\item \textbf{Phase 2 fails ($\eta > 0$ for AKLT)}: Incorrect SVD truncation logic or singular value sorting (descending order required).
		\item \textbf{Phase 3 fails (wrong environments)}: Missing complex conjugate on $A^{(i)*}$ or $B^{(i)*}$ in Eqs.~(4)-(5). Check environment indexing (off-by-one error $W^{(i)}$ vs $W^{(i\pm1)}$).
		\item \textbf{Phase 4 fails (wrong AKLT energy)}: MPO boundary conditions incorrect. Verify $b_{\text{left}}$ and $b_{\text{right}}$ select scalar component.
		\item \textbf{Phase 5 fails (stuck in local minimum)}: Insufficient bond dimension growth or missing noise injection (Section 4.5).
		\item \textbf{Phase 6 fails (variance $\sigma^2$ large)}: Lanczos solver tolerance too loose or $\mathcal{H}_{\text{eff}}$ matrix-vector product bug (check einsum indices).
	\end{itemize}
	
	\subsection{Reference Values for Heisenberg Chain}
	For the $S=1/2$ Heisenberg model with open boundary conditions, exact diagonalization values for small $L$:
	\begin{center}
		\begin{tabular}{@{}cc@{}}
			\toprule
			$L$ & $E_0 / J$ \\
			\midrule
			4 & $-1.616025403784...$ \\
			6 & $-2.493577131515...$ \\
			8 & $-3.374932109083...$ \\
			10 & $-4.258035965...$ \\
			\bottomrule
		\end{tabular}
	\end{center}
	Use these to verify MPO construction before attempting DMRG on larger systems.
	
	
\end{document}
