\documentclass[aps,prb,onecolumn,scriptaddress]{revtex4-2}
\usepackage{graphicx,amsmath,amsfonts,amssymb,bm,algorithm,algpseudocode,booktabs}
\usepackage[colorlinks=true, urlcolor=blue, linkcolor=blue, citecolor=blue]{hyperref}
\usepackage{tabularx}

\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\Id}{\mathbb{I}}
\newcommand{\eps}{\epsilon}
\newcommand{\Order}{\mathcal{O}}
\newcommand{\braket}[2]{\langle#1|#2\rangle}

\begin{document}
	
	\title{Numerical Implementation of MPS and DMRG}
	\author{Kimi 2.5}
	\author{Gemini 3.1}
	\affiliation{Based on the review article: U. Schollw\"ock, \textit{The density-matrix renormalization group in the age of matrix product states}, Annals of Physics 326 (2011) 96-192.}
	\date{\today}
	
	\begin{abstract}
		This document serves as a concrete numerical implementation guide for translating Matrix Product State (MPS) and Density Matrix Renormalization Group (DMRG) theory into robust, scalable code. It covers practical data structures, canonicalization, truncation, environment construction, explicit MPO formulations, and convergence diagnostics.
	\end{abstract}
	
	\maketitle
	
	\section{Data Structures}
	
	An MPS with open boundary conditions is stored as a list of 3-index tensors:
	\begin{itemize}
		\item $A^{(i)}_{a_{i-1},\sigma_i,a_i}$: Left-canonical tensor at site $i$, dimensions $(D_{i-1} \times d \times D_i)$
		\item $B^{(i)}_{a_{i-1},\sigma_i,a_i}$: Right-canonical tensor, dimensions $(D_{i-1} \times d \times D_i)$  
		\item $M^{(i)}_{a_{i-1},\sigma_i,a_i}$: General (non-canonical) tensor during optimization
		\item Physical dimension $d$ (e.g., $d=2$ for spin-1/2, $d=3$ for spin-1)
	\end{itemize}
	
	\textit{Storage Convention}: For site $i$, tensor $T^{(i)}$ has shape $(D_{i-1}, d, D_i)$ where the middle index is physical.
	
	\textit{Matrix Reshape Convention (Critical)}: When flattening $T^{(i)}_{a_{i-1},\sigma_i,a_i}$ to matrix $\tilde{M}_{r,c}$ for QR/SVD, the row index $r$ must follow C-order (row-major) mapping:
	\begin{equation}
		r = a_{i-1} \cdot d + \sigma_i \quad \text{where } a_{i-1} \in [0,D_{i-1}-1], \sigma_i \in [0,d-1]
	\end{equation}
	Column index $c = a_i$ (right bond). This yields matrix shape $(D_{i-1}d) \times D_i$.
	
	\textit{Python/NumPy Implementation}:
	\begin{verbatim}
		# C-order (row-major) reshape - CRITICAL for index consistency
		M_tilde = M.reshape((D_left * d, D_right), order='C')
		# Verify: M_tilde[a_prev * d + sigma, a_curr] == M[a_prev, sigma, a_curr]
	\end{verbatim}
	\textit{Fortran/MATLAB}: Use \texttt{reshape(..., order='F')} with reversed index ordering $r = \sigma_i \cdot D_{i-1} + a_{i-1}$ (column-major), or transpose after C-order reshape.
	
	\section{Canonicalization Algorithms}
	
	\subsection{Left-Canonicalization (QR-based)}
	More stable than SVD for canonicalization when singular values aren't needed.
	
	\begin{algorithm}[H]
		\caption{LeftCanonicalize (Numerically Stable)}
		\begin{algorithmic}[1]
			\Require MPS tensors $\{M^{(1)}, \dots, M^{(L)}\}$ with shapes $(D_{i-1}, d, D_i)$
			\Ensure Left-canonical $\{A^{(1)}, \dots, A^{(L)}\}$, last tensor $C$ (scalar or vector)
			\State $C \gets 1.0$ \Comment{Initial transfer matrix, $1 \times 1$}
			\For{$i = 1$ to $L-1$}
			\State \Comment{Contract incoming transfer matrix $C$ with current site}
			\State $\tilde{M}_{(a_{i-1}\sigma_i), a_i} \gets \sum_{a'_{i-1}} C_{a_{i-1}, a'_{i-1}} M^{(i)}_{a'_{i-1}, \sigma_i, a_i}$ 
			\State \Comment{For $i=1$: $C$ is $1\times 1$, so $\tilde{M}$ is simply $M^{(1)}$ reshaped to $(d) \times D_1$}
			\State \Comment{Matrix shape: $(D_{i-1}d) \times D_i$}
			\State \Comment{QR decomposition with column-pivoting for stability}
			\If{$D_{i-1} \cdot d < D_i$}
			\State \textbf{Warning:} Bond dimension $D_i$ exceeds physical capacity $D_{i-1}d$. 
			\State \phantom{\textbf{Warning:}} State is necessarily rank-deficient. Perform truncation or use random initialization.
			\EndIf
			\State $D_{\text{new}} \gets \min(D_{i-1} \cdot d, D_i)$
			\State $Q, R \gets \text{QR}(\tilde{M})$ \Comment{Economic QR. $Q$: $(D_{i-1}d) \times D_{\text{new}}$, $R$: $D_{\text{new}} \times D_i$}
			\State $A^{(i)}_{a_{i-1}, \sigma_i, a_i} \gets \text{reshape}(Q, (D_{i-1}, d, D_{\text{new}}))$
			\State $C \gets R$ \Comment{Transfer matrix to next site, shape $D_{\text{new}} \times D_i$}
			\EndFor
			\State \Comment{Final site: absorb last $C$}
			\State $A^{(L)}_{a_{L-1}, \sigma_L, 1} \gets \sum_{a'_{L-1}} C_{a_{L-1}, a'_{L-1}} M^{(L)}_{a'_{L-1}, \sigma_L, 1}$ \Comment{Shape: $(D_{L-1}, d, 1)$}
		\end{algorithmic}
	\end{algorithm}
	
	
	\subsection{Mixed-Canonical Form at Bond $k$}
	Required for variational minimization.
	
	\begin{algorithm}[H]
		\caption{BringToBond\\
			\textit{Numerical Note:} RQ decomposition (used for right-canonicalization) is less stable than QR and may not be available in all libraries (e.g., PyTorch). Alternative: Reverse the site order, apply \textsc{LeftCanonicalize}, then reverse back.}
		\begin{algorithmic}[1]
			\Require MPS $\{M^{(i)}\}$, target bond $k$
			\Ensure Left-canonical $A^{(1..k)}$, right-canonical $B^{(k+1..L)}$, center matrix $S$
			\State Left-canonicalize sites $1 \dots k$ using QR sweeps
			\State Right-canonicalize sites $L \dots k+1$ using RQ sweeps (reverse direction)
			\State $S \gets$ matrix at bond $k$ (contains singular values if from SVD)
		\end{algorithmic}
	\end{algorithm}
	
	\section{Truncation and Compression}
	
	\subsection{SVD Compression with Tolerance}
	\begin{algorithm}[H]
		\caption{TruncateBond}
		\begin{algorithmic}[1]
			\Require Matrix $M$ (dimensions $m \times n$), max bond $D_{\max}$, tolerance $\eps$, \textbf{abs tolerance $\eps_{\text{abs}} \sim 10^{-14}$}
			\Ensure Truncated matrices $U, S, V^\dagger$ with $D \leq D_{\max}$
			\State $U, S, V^\dagger \gets \text{SVD}(M, \text{full\_matrices=False})$
			\State $s_i \gets \text{diag}(S)$ in descending order
			\State \Comment{Determine effective $D$ using relative and absolute tolerances}
			\State \Comment{Safety check: prevent $D=0$ when all singular values vanish}
			\If{$s_0 < \eps_{\text{abs}}$}
			\State \textbf{Warning:} All singular values below absolute tolerance.
			\State Set $D \gets 1$ (keep single largest, even if numerically zero).
			\Else
			\State $D \gets \min\left(D_{\max}, \text{count}\left(s_i > \max\left(\eps \cdot s_0, \eps_{\text{abs}}\right)\right)\right)$
			\State \textbf{where} $\eps_{\text{abs}} \sim 10^{-14}$ (double precision) prevents $D \to \infty$ when $s_0 \approx 0$
			\EndIf
			\State \Comment{Truncation error}
			\State $\eta \gets \sum_{i=D}^{r-1} s_i^2$ \Comment{Discarded weight}
			\State $U \gets U_{:,0:D-1}$, $S \gets S_{0:D-1,0:D-1}$, $V^\dagger \gets V^\dagger_{0:D-1,:}$
			\State \Return $U, S, V^\dagger, \eta$
		\end{algorithmic}
	\end{algorithm}
	
	\textit{Numerical Note}: For $\eps \sim 10^{-12}$, single-precision floats may underflow. Use double precision (complex128).
	
	\section{Ground State Algorithm: Single-Site DMRG}
	
	\subsection{Effective Environment Construction}
	The local Hamiltonian at site $i$ is built from pre-computed environments.
	For numerical implementation, we establish a strict universal index convention to prevent shape mismatches and broadcasting errors:
	\begin{itemize}
		\item \textbf{MPO bonds}: \texttt{b} (left), \texttt{B} (right)
		\item \textbf{MPS bra bonds}: \texttt{x} (left), \texttt{X} (right)
		\item \textbf{MPS ket bonds}: \texttt{y} (left), \texttt{Y} (right)
		\item \textbf{Physical indices}: \texttt{s} (bra/out), \texttt{t} (ket/in)
	\end{itemize}
	
	The environments are always stored as shape \texttt{(MPO\_bond, Bra\_bond, Ket\_bond)}. The MPO is stored as \texttt{(Left, Right, Out, In)}. In Python, the left environment update $L^{(i)}$ is computed as:
	\begin{verbatim}
		# Left Environment Update (site i-1 -> i)
		# L_old shape: (b, x, y)
		# A_ket shape: (y, t, Y) -> The actual MPS tensor at site i-1
		# W shape:     (b, B, s, t) -> The MPO tensor at site i-1
		# A_bra shape: (x, s, X) -> The complex conjugate of A_ket
		# Output L_new shape: (B, X, Y)
		
		L_new = np.einsum('bxy, ytY, bBst, xsX -> BXY', 
		L_old, A_ket, W, A_ket.conj(), optimize=True)
	\end{verbatim}
	\textit{Note on Performance}: Always use \texttt{optimize=True} in NumPy's einsum, which automatically finds the optimal contraction path, mimicking the $\Order(D^3)$ scaling of step-by-step tensor network contractions.
	
	Similarly, the right environment update $R^{(i)}$ (constructed backwards from $L$ down to 1) is:
	\begin{verbatim}
		# Right Environment Update (site i+1 -> i)
		# R_old shape: (B, X, Y)
		# B_ket shape: (y, t, Y)
		# W shape:     (b, B, s, t)
		# B_bra shape: (x, s, X)
		# Output R_new shape: (b, x, y)
		
		R_new = np.einsum('BXY, ytY, bBst, xsX -> bxy', 
		R_old, B_ket, W, B_ket.conj(), optimize=True)
	\end{verbatim}
	
	\textit{Verification}: After each update, check environment Hermiticity: $\|L - L^\dagger\|_{\max} < 10^{-14}$.
	
	\textit{Index Dependency Verification}:
	\begin{itemize}
		\item $L^{(i)}$ depends on sites $1 \dots i-1$, hence uses $W^{(i-1)}$ (not $W^{(i)}$).
		\item $R^{(i)}$ depends on sites $i+1 \dots L$, hence uses $W^{(i+1)}$ (not $W^{(i)}$).
		\item \textbf{Common bug}: Off-by-one error in MPO index. Verify: $L^{(2)}$ uses $W^{(1)}$, $R^{(L-1)}$ uses $W^{(L)}$.
	\end{itemize}
	
	\subsection{Local Optimization (Matrix-Free Formulation)}
	Solve the eigenvalue problem for the center tensor $M^{(i)}_{\alpha_{i-1},\alpha_i}$ reshaped as vector $v$ of length $d D_{i-1} D_i$:
	
	\begin{equation}
		\mathcal{H}_{\text{eff}} \, v = E v
	\end{equation}
	
	\textbf{Scalability Note}: Explicitly constructing the matrix $\mathcal{H}_{\text{eff}}$ requires $\Order(d^2 D^4)$ memory and destroys scalability. Therefore, the action of $\mathcal{H}_{\text{eff}}$ on the state vector $v$ must be computed \textit{matrix-free} via sequential tensor contractions. In Python, this is accomplished by wrapping Algorithm 4 in a \texttt{scipy.sparse.linalg.LinearOperator} and passing it to \texttt{eigsh}. This bypasses the $D^4$ bottleneck, reducing memory to $\Order(D^2)$ and computation time to $\Order(D^3)$.
	
	\textit{Lanczos stopping criterion}: Stop when eigenvalue residual $\|\mathcal{H}_{\text{eff}}v - Ev\| < \delta$ where $\delta \sim 10^{-10}$ or when $N_{\text{iter}} > 20-50$.
	
	\begin{algorithm}[H]
		\caption{ApplyEffectiveHamiltonian (Matrix-Free)}
		\begin{algorithmic}[1]
			\Require Vector $v$ (flattened from $V_{y,t,Y}$), $L$ shape $(b, x, y)$, $R$ shape $(B, X, Y)$, $W$ shape $(b, B, s, t)$
			\Ensure $(\mathcal{H}_{\text{eff}}v)_{x,s,X}$ as flattened vector (maps ket to bra space)
			\State Reshape $v \to V_{y,t,Y}$ \Comment{Shape $(D_{i-1}, d, D_i)$}
			\State \textbf{Step 1 (Contract Left)}: $T_1 = \sum_{y} L_{b, x, y} V_{y, t, Y}$
			\State \phantom{\textbf{Step 1}} \textit{Python}: \texttt{T1 = np.einsum('bxy, ytY -> bxtY', L, V, optimize=True)}
			\State \textbf{Step 2 (Contract MPO)}: $T_2 = \sum_{b, t} T_{1(b, x, t, Y)} W_{b, B, s, t}$  
			\State \phantom{\textbf{Step 2}} \textit{Python}: \texttt{T2 = np.einsum('bxtY, bBst -> BxsY', T1, W, optimize=True)} 
			\State \textbf{Step 3 (Contract Right)}: $V_{\text{new}} = \sum_{B, Y} T_{2(B, x, s, Y)} R_{B, X, Y}$
			\State \phantom{\textbf{Step 3}} \textit{Python}: \texttt{V\_new = np.einsum('BxsY, BXY -> xsX', T2, R, optimize=True)} 
			\State \Return \texttt{V\_new.flatten()} \Comment{Pass back to Lanczos/Davidson solver}
		\end{algorithmic}
	\end{algorithm}
	\textit{Cost}: $\Order(D_W D_{i-1} D_i d^2 + D_W^2 D_{i-1} D_i d)$. 
	
	\textit{Numerical Note on Hermiticity}: Due to floating-point accumulation during sweeps, the effective Hamiltonian may lose exact machine-precision Hermiticity. This often causes iterative solvers like \texttt{scipy.sparse.linalg.eigsh} to fail. However, you must \textbf{not} explicitly symmetrize the individual environments $L$ and $R$ (e.g., $L \gets \frac{1}{2}(L + L^\dagger)$). The individual MPO bond channels (like $\hat{S}^+$) are strictly non-Hermitian; symmetrizing them corrupts the Hamiltonian. Instead, simply pass \texttt{ishermitian=True} (or explicitly provide \texttt{rmatvec=matvec}) when defining the \texttt{LinearOperator} so SciPy mathematically enforces Hermiticity internally during Lanczos iterations.
	
	\subsection{Environment Initialization}
	Before sweeping, initialize boundary environments consistent with the MPO structure:
	\begin{align}
		L^{(1)}_{b_0}(a_0, a'_0) &= \delta_{b_0, b_{\text{left}}} \delta_{a_0, a'_0} & \text{(Left boundary: $1 \times 1$ matrices)} \\
		R^{(L)}_{b_L}(a_L, a'_L) &= \delta_{b_L, b_{\text{right}}} \delta_{a_L, a'_L} & \text{(Right boundary: $1 \times 1$ matrices)}
	\end{align}
	where $b_{\text{left}}$ and $b_{\text{right}}$ are the MPO indices selecting the Hamiltonian scalar. In code, this means $L^{(1)}$ is a $(D_W, 1, 1)$ tensor with a $1.0$ at index $b_{\text{left}}$ and $0$ elsewhere.
	
	\subsection{Complex Arithmetic Handling}
	Even for real Hamiltonians (e.g., Heisenberg), maintain \texttt{complex128} dtype throughout:
	\begin{itemize}
		\item \textbf{Left environment:} Uses complex conjugate only (no transpose): Python: \texttt{A.conj()}, \textbf{not} \texttt{A.T.conj()}.
		\item \textbf{Right environment:} Uses complex conjugate only when performing the explicit \texttt{np.einsum} specified above.
		\item \textbf{Verification:} The density matrix $\rho = M M^\dagger$ must be Hermitian. Check $\|\rho - \rho^\dagger\|_{\max} < 10^{-14}$.
	\end{itemize}
	
	\subsection{Datatype and Library Requirements}
	
	\textit{Floating Point Precision}: All tensors must use \texttt{complex128} (double precision). Single precision (\texttt{float32}) leads to underflow in $\epsilon \sim 10^{-12}$ truncations.
	
	\textit{QR Decomposition}: Standard \texttt{numpy.linalg.qr} lacks column-pivoting. Use:
	\begin{verbatim}
		from scipy.linalg import qr
		Q, R, P = qr(M, pivoting=True, mode='economic')
		# Reorder Q according to permutation P if necessary
	\end{verbatim}
	Ensure diagonal elements of $R$ have non-negative real parts (multiply column $j$ of $Q$ by $\text{sign}(R_{jj})$ if needed).
	
	\textit{SVD Truncation}: Explicitly sort singular values descending:
	\begin{verbatim}
		U, s, Vh = np.linalg.svd(M, full_matrices=False)
		# Ensure descending (NumPy default) - reverse if using different library
		idx = np.argsort(s)[::-1]
		s, U, Vh = s[idx], U[:, idx], Vh[idx, :]
	\end{verbatim}
	
	\subsection{Sweep Algorithm}
	
	\begin{algorithm}[H]
		\caption{SingleSiteDMRG (with Gauge Control)}
		\begin{algorithmic}[1]
			\Require Hamiltonian MPO $\{W^{(i)}\}$, initial MPS guess $\{M^{(i)}\}$, max bond $D_{\max}$, tolerance $\eps$
			\Ensure Ground state energy $E_0$ and optimized MPS in mixed-canonical form
			\State Right-canonicalize initial MPS using RQ sweeps (or reverse QR)
			\State Initialize all $R^{(i)}$ environments via backward contraction from site $L$
			\State $L^{(1)} \gets$ left boundary tensor (see initialization above)
			\While{not converged}
			\State \textbf{Right sweep} ($i=1$ to $L-1$):
			\For{$i=1$ to $L-1$}
			\State Build $\mathcal{H}_{\text{eff}}$ from $L^{(i)}$, $R^{(i)}$, $W^{(i)}$
			\State Solve $\mathcal{H}_{\text{eff}} v = E v$ using \textsc{ApplyEffectiveHamiltonian} and Lanczos
			\State Reshape $v \to M^{(i)}_{a_{i-1},\sigma_i,a_i}$
			\State \Comment{Gauge update: SVD and shift center to the right}
			\State Reshape $M^{(i)} \to \tilde{M}_{(a_{i-1}\sigma_i), a_i}$ (matrix $(dD_{i-1}) \times D_i$)
			\State $U, S, V^\dagger \gets \text{SVD}(\tilde{M})$, truncate to $D_{\max}$
			\State $D_{\text{new}} \gets \min(D_{\max}, \text{len}(S))$
			\State $A^{(i)} \gets \text{reshape}(U, (D_{i-1}, d, D_{\text{new}}))$ \Comment{Left-canonical}
			\State \Comment{Absorb singular values into next site (center moves right)}
			\State $M^{(i+1)}_{a_i,\sigma_{i+1},a_{i+2}} \gets \sum_{a'_i, a''_i} \text{diag}(S)_{a_i,a'_i} V^\dagger_{a'_i,a''_i} M^{(i+1)}_{a''_i,\sigma_{i+1},a_{i+2}}$
			\State Update $L^{(i+1)}$ using $A^{(i)}$ and $W^{(i)}$
			\EndFor
			\State \textbf{Left sweep} ($i=L$ to $2$):
			\For{$i=L$ to $2$}
			\State Solve eigenproblem using updated environments
			\State Reshape $M^{(i)} \to \tilde{M}_{a_{i-1}, (\sigma_i a_i)}$ (matrix $D_{i-1} \times (dD_i)$)
			\State $U, S, V^\dagger \gets \text{SVD}(\tilde{M})$, truncate to $D_{\max}$
			\State $D_{\text{new}} \gets \min(D_{\max}, \text{len}(S))$
			\State $B^{(i)} \gets \text{reshape}(V^\dagger, (D_{\text{new}}, d, D_i))$ \Comment{Right-canonical}
			\State \Comment{Absorb singular values into previous site (center moves left)}
			\State $M^{(i-1)}_{a_{i-2},\sigma_{i-1},a_{i-1}} \gets \sum_{a'_{i-1}, a''_{i-1}} M^{(i-1)}_{a_{i-2},\sigma_{i-1},a''_{i-1}} U_{a''_{i-1},a'_{i-1}} \text{diag}(S)_{a'_{i-1},a_{i-1}}$
			\State Update $R^{(i-1)}$ using $B^{(i)}$ and $W^{(i)}$
			\EndFor
			\State Check convergence (ALL must hold):
			\State \quad 1. Energy: $|\Delta E| = |E_n - E_{n-1}| < \eps_E$ (e.g., $10^{-8}$ per site)
			\State \quad 2. \textbf{Variance}: $\sigma^2 = \bra{\psi}H^2\ket{\psi} - E^2 < \eps_{\sigma}$ (stricter, $\sim 10^{-6}$)
			\State \quad 3. \textbf{State convergence}: $\max_i \|A^{(i)}_{\text{new}} - A^{(i)}_{\text{old}}\|_{\text{Frob}} < \eps_{\text{tensor}}$
			\EndWhile
		\end{algorithmic}
	\end{algorithm}
	
	\textit{Critical Implementation Note — The Local Minimum Problem}:
	Single-site updates with fixed bond dimension $D$ cannot escape local minima during phase transitions or for systems with symmetries. Two mitigation strategies:
	\begin{enumerate}
		\item \textbf{Two-Site DMRG}: Use Algorithm 6 for the first 3-5 sweeps to dynamically increase $D$, then switch to single-site for speed.
		\item \textbf{Perturbation/Noise Injection}: Before truncation, add random noise to the singular values or density matrix.
	\end{enumerate}
	
	\section{Two-Site DMRG (Robust against local minima)}
	
	For systems with symmetries or near phase transitions, single-site DMRG may get stuck. Two-site updates allow bond dimension adaptation.
	
	\textit{Critical Initialization Note:} When initializing the random MPS for Two-Site DMRG, do not artificially restrict the initial state to a small bond dimension (e.g., $D=2$) with the expectation that it will organically grow. For Hamiltonians with strict symmetries (like $SU(2)$ Heisenberg), a small random state may become trapped in a specific symmetry sector. In such cases, the local Effective Hamiltonian has zero matrix elements mapping to other sectors, causing the SVD to yield exact zeroes for new singular values, completely preventing $D$ from growing. Always initialize the random starting tensors to the full target capacity $D_{\max}$ to ensure the algorithm can explore the full Hilbert space.
	
	\begin{algorithm}[H]
		\caption{TwoSiteUpdate}
		\begin{algorithmic}[1]
			\Statex \textit{Effective Hamiltonian on two-site space:}
			\Statex $\mathcal{H}_{\text{eff}}^{(i,i+1)} = \sum_{b_{i-1}, b_{i+1}} L_{b_{i-1}} \otimes W^{(i)}_{b_{i-1},b_i} \otimes W^{(i+1)}_{b_i,b_{i+1}} \otimes R_{b_{i+1}}$
			\Statex Dimensions: $d^2 D_{i-1} D_{i+1} \times d^2 D_{i-1} D_{i+1}$.
			\Require Current bond dimensions $D_{i-1}, D_i$, max bond $D_{\max}$, tolerance $\eps$					
			\State Contract: $\Theta_{\alpha_{i-1},\sigma_i,\sigma_{i+1},\alpha_{i+1}} = \sum_{\alpha_i} A^{(i)}_{\alpha_{i-1},\sigma_i,\alpha_i} B^{(i+1)}_{\alpha_i,\sigma_{i+1},\alpha_{i+1}}$
			\State Reshape: $\Theta \to \Theta_{(\alpha_{i-1}\sigma_i), (\sigma_{i+1}\alpha_{i+1})}$ (matrix of size $(d D_{i-1}) \times (d D_{i+1})$)
			\State Build $\mathcal{H}_{\text{eff}}$ on combined two-site space using sequential contractions.
			\State Solve for lowest eigenvector $\tilde{\Theta}$ (size $d^2 D_{i-1} D_{i+1}$) using Lanczos/Davidson
			\State Reshape $\tilde{\Theta} \to$ matrix $(d D_{i-1}) \times (d D_{i+1})$
			\State SVD: $\tilde{\Theta} = U S V^\dagger$
			\State Truncate to $D \leq D_{\max}$ singular values using \textsc{TruncateBond}
			\State $A^{(i)} \gets \text{reshape}(U, (D_{i-1}, d, D))$ \Comment{Left-canonical}
			\State $B^{(i+1)} \gets \text{reshape}(S V^\dagger, (D, d, D_{i+1}))$ \Comment{Right-canonical}
			\State \Return Updated tensors, truncation error $\eta = \sum_{\text{discarded}} s_i^2$
		\end{algorithmic}
	\end{algorithm}
	
	\section{MPO Construction for Hamiltonians}
	
	\subsection{From Operators to 4D Tensors}
	\label{sec:mpo_theory}
	In the literature, MPOs are often written as matrices of operators $\hat{W}_{b_{i-1}, b_i}$. To translate this into a 4D array $W^{(i)}$ of shape $(D_W, D_W, d, d)$ required for coding, one evaluates the matrix elements in the local physical basis $\ket{\sigma}$:
	\begin{equation}
		W^{(i)}_{b_{i-1}, b_i, \sigma, \sigma'} = \bra{\sigma} \hat{W}_{b_{i-1}, b_i} \ket{\sigma'}
	\end{equation}
	For practical implementation, define the local spin matrices explicitly. For spin-1/2 ($d=2$):
	\begin{equation}
		S^+ = \begin{pmatrix} 0 & 1 \\ 0 & 0 \end{pmatrix}, \quad 
		S^- = \begin{pmatrix} 0 & 0 \\ 1 & 0 \end{pmatrix}, \quad 
		S^z = \begin{pmatrix} 0.5 & 0 \\ 0 & -0.5 \end{pmatrix}, \quad 
		\Id = \begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix}
	\end{equation}
	For spin-1 ($d=3$):
	\begin{equation}
		S^x = \frac{1}{\sqrt{2}} \begin{pmatrix} 0 & 1 & 0 \\ 1 & 0 & 1 \\ 0 & 1 & 0 \end{pmatrix}, \quad 
		S^y = \frac{i}{\sqrt{2}} \begin{pmatrix} 0 & -1 & 0 \\ 1 & 0 & -1 \\ 0 & 1 & 0 \end{pmatrix}, \quad 
		S^z = \begin{pmatrix} 1 & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & -1 \end{pmatrix}
	\end{equation}
	
	\subsection{Nearest-Neighbor Heisenberg Model}
	\label{sec:mpo_heis}
	MPO bond dimension $D_W = 5$. Define operator-valued matrices $\hat{W}^{(i)}$:
	
	\begin{equation}
		\hat{W}^{(i)} = \begin{pmatrix}
			\hat{\Id} & \cdot & \cdot & \cdot & \cdot \\
			\hat{S}^+ & \cdot & \cdot & \cdot & \cdot \\
			\hat{S}^- & \cdot & \cdot & \cdot & \cdot \\
			\hat{S}^z & \cdot & \cdot & \cdot & \cdot \\
			-h\hat{S}^z & \frac{J}{2}\hat{S}^- & \frac{J}{2}\hat{S}^+ & J_z \hat{S}^z & \hat{\Id}
		\end{pmatrix}
	\end{equation}
	
	Boundary conditions (0-indexed array convention):
	\begin{itemize}
		\item Site 1 (left boundary): $W^{(1)}$ is a $(1, D_W, d, d)$ tensor selecting the last row (index 4 for $D_W=5$). \\
		\textbf{Code}: \texttt{W\_left[0, :, :, :] = bulk\_W[4, :, :, :]}
		
		\item Site $L$ (right boundary): $W^{(L)}$ is a $(D_W, 1, d, d)$ tensor selecting the first column (index 0). \\
		\textbf{Code}: \texttt{W\_right[:, 0, :, :] = bulk\_W[:, 0, :, :]}
	\end{itemize}
	
	\textit{Index Verification}: MPO bond indices $b_i \in \{0, 1, \dots, D_W-1\}$. For Heisenberg with $D_W=5$:
	\begin{itemize}
		\item $b_{\text{left}} = 4$ (index of $\hat{\Id}$ in last row)
		\item $b_{\text{right}} = 0$ (index of $\hat{\Id}$ in first column)
	\end{itemize}
	Boundary environments: $L^{(1)}_{b_0} = \delta_{b_0, 4}$, $R^{(L)}_{b_L} = \delta_{b_L, 0}$.
	
	\subsection{AKLT Model}
	\label{sec:mpo_aklt}
	The AKLT model is a spin-1 ($d=3$) system involving a biquadratic interaction:
	\begin{equation}
		H_{\text{AKLT}} = \sum_{i=1}^{L-1} \left[ \vec{S}_i \cdot \vec{S}_{i+1} + \frac{1}{3} (\vec{S}_i \cdot \vec{S}_{i+1})^2 \right]
	\end{equation}
	To construct a highly scalable exact MPO, we cast the interaction as a sum of outer products. Define the 3-component vector of spin operators $\mathbf{S} = (S^x, S^y, S^z)^T$. Define $\mathbf{Q}$ as the 9-component vector containing all pairwise products $Q^{ab} = S^a S^b$ for $a,b \in \{x,y,z\}$. The nearest-neighbor interaction perfectly decouples into $\mathbf{S}_i \cdot \mathbf{S}_{i+1} + \frac{1}{3} \mathbf{Q}_i \cdot \mathbf{Q}_{i+1}$.
	
	The resulting MPO tensor $\hat{W}^{(i)}$ is elegantly expressed as a block matrix with bond dimension $D_W = 1 + 3 + 9 + 1 = 14$:
	\begin{equation}
		\hat{W}^{(i)} = \begin{pmatrix}
			\hat{\Id} & \mathbf{0}_{1 \times 3} & \mathbf{0}_{1 \times 9} & 0 \\
			\mathbf{S} & \mathbf{0}_{3 \times 3} & \mathbf{0}_{3 \times 9} & \mathbf{0}_{3 \times 1} \\
			\mathbf{Q} & \mathbf{0}_{9 \times 3} & \mathbf{0}_{9 \times 9} & \mathbf{0}_{9 \times 1} \\
			0 & \mathbf{S}^T & \frac{1}{3}\mathbf{Q}^T & \hat{\Id}
		\end{pmatrix}
	\end{equation}
	\textit{Boundary Initialization}: For the left boundary $W^{(1)}$, select the bottom row ($b_{\text{left}} = 13$). For the right boundary $W^{(L)}$, select the first column ($b_{\text{right}} = 0$). \\
	\textbf{Code}: \texttt{W\_left[0, :, :, :] = bulk\_W[13, :, :, :]} and \texttt{W\_right[:, 0, :, :] = bulk\_W[:, 0, :, :]}.
	
	
	\section{Compression of MPO-MPS Product}
	When applying MPO to MPS, bond dimension grows to $D \cdot D_W$. Compress variationally:
	
	\begin{algorithm}[H]
		\caption{VariationalMPSCompression}
		\begin{algorithmic}[1]
			\Require Target MPS $\ket{\psi}$ (bond $D_{\psi}$), initial guess $\ket{\phi}$ (bond $D_{\phi} < D_{\psi}$)
			\Ensure $\ket{\phi} \approx \ket{\psi}$ with minimal $\|\phi - \psi\|^2$
			\State Bring $\ket{\phi}$ to mixed-canonical form at bond $k=1$ (right-canonicalized)
			\State Compute initial environments $L_{\psi\phi}, R_{\psi\phi}$ (overlaps between $\psi$ and $\phi$)
			\While{not converged}
			\For{$i=1$ to $L-1$} \Comment{Right sweep}
			\State $T_{a_{i-1},\sigma_i,a_i} \gets \sum_{\{a'\}} L_{\psi\phi} \cdot M^{(i)}_{\psi} \cdot R_{\psi\phi}$
			\If{$\ket{\phi}$ is kept strictly in mixed-canonical form at site $i$}
			\State $M^{(i)}_{\phi} \gets T_{a_{i-1},\sigma_i,a_i}$ \Comment{Metric $N=\Id$, avoids costly matrix inversion!}
			\Else
			\State $N_{a_{i-1},a_i} \gets \sum_{b_i} L^{(i)}_{b_i} R^{(i)}_{b_i}$ \Comment{Metric, shape $D_{\phi} \times D_{\phi}$}
			\State $M^{(i)}_{\phi} \gets \text{Reshape}(N^{-1} \cdot \text{Flatten}(T))$ \Comment{Pseudo-inverse if singular}
			\EndIf
			\State SVD and shift gauge right (as in DMRG)
			\EndFor
			\State Left sweep (analogous)
			\EndWhile
		\end{algorithmic}
	\end{algorithm}
	
	\section{Model Examples and Physics}
	
	\subsection{Spin-1/2 Heisenberg Chain}
	The ground state exhibits quasi-long-range antiferromagnetic order (power-law decay of correlations $\sim (-1)^r/r$ at criticality). 
	
	\textbf{Numerical Notes:} The Heisenberg ground state is not exactly representable by a finite-$D$ MPS. Due to its critical nature, the entanglement entropy scales logarithmically with system size, requiring bond dimensions $D \sim L^{\alpha}$ (with $\alpha \approx 1/6$) to maintain accuracy. Results must be extrapolated in the truncation error $\eta$ or $1/D$.
	
	\subsection{AKLT Model (Exact MPS Representation)}
	Unlike the Heisenberg model, the AKLT ground state is gapped and can be represented \textit{exactly} as a translationally invariant MPS with bond dimension $D=2$.
	
	\textbf{Explicit MPS Matrices:} The left-canonical matrices $A^{\sigma}$ (where $\sigma \in \{+,0,-\}$ corresponds to $m_z = +1, 0, -1$) are given by
	\begin{align}
		A^{+} &= \begin{pmatrix} 0 & \sqrt{2/3} \\ 0 & 0 \end{pmatrix}, \\
		A^{0} &= \begin{pmatrix} -1/\sqrt{3} & 0 \\ 0 & 1/\sqrt{3} \end{pmatrix}, \\
		A^{-} &= \begin{pmatrix} 0 & 0 \\ -\sqrt{2/3} & 0 \end{pmatrix}.
	\end{align}
	These satisfy the left-normalization condition $\sum_{\sigma} A^{\sigma\dagger} A^{\sigma} = \mathbb{I}_2$.
	
	\textit{Verification Test}: After implementing the MPS class, verify:
	\begin{equation}
		\sum_{\sigma \in \{+,0,-\}} A^{\sigma\dagger} A^{\sigma} = \mathbb{I}_2, \quad 
		\sum_{\sigma} A^{\sigma} A^{\sigma\dagger} = \frac{2}{3}\mathbb{I}_2 + \frac{1}{3}\sigma_z \neq \mathbb{I}
	\end{equation}
	The left-canonical condition is satisfied; the right-canonical is not (this is an \textit{injective} MPS). For two-site DMRG initialization with $D=2$ and the MPO provided in Sec.~\ref{sec:mpo_aklt}, the energy should converge to exactly $E_0 = -(L-1)\cdot \frac{2}{3}$ with truncation error $\eta \equiv 0$.
	
	\textbf{Properties:}
	\begin{itemize}
		\item \textbf{Correlation Functions:} Spin-spin correlations decay exponentially as $\langle S^z_i S^z_j \rangle \sim (-1/3)^{|i-j|}$ with correlation length $\xi = 1/\ln 3 \approx 0.91$.
		\item \textbf{Hidden Order:} The model exhibits a non-local "string order" $\langle S^z_i \exp(i\pi \sum_{k=i+1}^{j-1} S^z_k) S^z_j \rangle = -4/9$ for $|i-j|>2$, indicating topological order.
		\item \textbf{Open Boundaries:} For OBC, the exact ground state is $\ket{\psi} = \sum_{\{\sigma\}} A^{\sigma_1} \cdots A^{\sigma_L} \ket{\sigma_1 \cdots \sigma_L}$, resulting in two free $S=1/2$ edge spins.
	\end{itemize}
	
	This model serves as an excellent benchmark for DMRG implementations: with $D=2$, the algorithm should reproduce the exact energy $E_0 = -(L-1) \cdot 2/3$ (for $J=1$) and exponentially decaying correlations without truncation error.
	
	\section{Convergence Diagnostics}
	
	\begin{itemize}
		\item \textbf{Energy convergence}: $\Delta E = |E_{n} - E_{n-1}| < 10^{-8}$ (per site)
		\item \textbf{Truncation error}: Monitor $\eta = \sum_{\text{discarded}} s_i^2$. For accurate results, extrapolate $E(\eta) \to \eta \to 0$.
		\item \textbf{Variance}: $\sigma^2 = \bra{\psi}H^2\ket{\psi} - \bra{\psi}H\ket{\psi}^2$. Should be $\sim D^{-2}$ for gapped systems.
		\item \textbf{Entanglement entropy}: Check $S = -\sum s_i^2 \ln s_i^2$ saturation across bonds.
	\end{itemize}
	
	\textit{Computational Note on Variance:} Computing $\langle H^2 \rangle$ by squaring the MPO 
	(bond dimension $D_W^2$) is inefficient. Instead:
	\begin{enumerate}
		\item Compute $\ket{\phi} = H\ket{\psi}$ via MPO-MPS multiplication (cost $\sim LD_WD^3$).
		\item Compute $\langle H^2 \rangle = \braket{\phi}{\phi}$ (MPS overlap).
		\item Reuse $\langle\psi|H|\psi\rangle$ from DMRG energy for $E^2$ term.
	\end{enumerate}
	This avoids the $D_W^2$ scaling of explicit $H^2$ construction.
	
	\section{Practical Implementation Notes (Python/NumPy)}
	\label{sec:implementation_notes}
	
	While the mathematical algorithms provide the rigorous framework for DMRG, translating them into high-performance Python code requires navigating several realities of array programming and memory management. The following details are critical for a robust, production-ready codebase.
	
	\subsection{Eigensolver Initial Guess (\texttt{v0})}
	In Algorithm 5, solving the effective Hamiltonian is the primary computational bottleneck. When using iterative Lanczos/Davidson solvers like \texttt{scipy.sparse.linalg.eigsh}, you \textbf{must} pass the tensor from the previous sweep as the initial guess vector. If omitted, SciPy initializes a random vector, increasing the number of iterations (and thus computation time) by an order of magnitude.
	
	\begin{verbatim}
		from scipy.sparse.linalg import LinearOperator, eigsh
		
		# Define the matrix-vector product wrapper
		def matvec(v):
		# Reshape v, apply contractions (Algorithm 4), return flattened result
		pass
		
		Heff = LinearOperator((dim, dim), matvec=matvec)
		
		# CRITICAL: Pass the current flattened tensor as the initial guess 'v0'
		evals, evecs = eigsh(Heff, k=1, which='SA', v0=M_current.flatten(), tol=1e-10)
		M_optimized = evecs[:, 0].reshape(shape, order='C')
	\end{verbatim}
	
	\subsection{Gauge Shifting via \texttt{tensordot}}
	In the gauge update step of DMRG (absorbing singular values into the adjacent site), the operation is mathematically $M^{(i+1)} = \sum \text{diag}(S) \cdot V^\dagger \cdot M^{(i+1)}$. Do not use \texttt{np.einsum} for this matrix-tensor multiplication. Instead, use \texttt{np.tensordot}, which maps directly to highly optimized BLAS \texttt{dgemm}/\texttt{zgemm} routines.
	
	\begin{verbatim}
		# 1. Form the gauge matrix to push right
		gauge_matrix = np.diag(S) @ Vh  # Shape: (D_new, D_right)
		
		# 2. Push into the next site using tensordot
		# axes=([1], [0]) sums over the 2nd index of gauge and 1st index of M_next
		M_next = np.tensordot(gauge_matrix, M_next, axes=([1], [0])) 
	\end{verbatim}
	
	\subsection{Environment Caching}
	To maintain the strict $\Order(1)$ per-site scaling of the sweep algorithm (excluding the internal $D^3$ tensor contractions), environments must not be recomputed from scratch. Maintain two caches (e.g., Python lists) for the left and right environments. 
	
	\begin{verbatim}
		# Initialize caches. L_env[i] holds the left environment acting ON site i.
		L_env = [None] * L
		R_env = [None] * L
		
		# Initialize boundaries (0-indexed)
		L_env[0] = np.zeros((D_w, 1, 1), dtype=np.complex128)
		L_env[0][b_left, 0, 0] = 1.0
		
		R_env[L-1] = np.zeros((D_w, 1, 1), dtype=np.complex128)
		R_env[L-1][b_right, 0, 0] = 1.0
		
		# Pre-calculate R_env for the initial right-to-left sweep
		for i in range(L-1, 0, -1):
		R_env[i-1] = update_right_env(R_env[i], B[i], W[i])
	\end{verbatim}
	When moving right (from site $i$ to $i+1$), calculate the new environment using \texttt{L\_env[i]} and save the result directly to \texttt{L\_env[i+1]}.
	
	\subsection{Memory Views vs. Copies in \texttt{reshape}}
	In NumPy, \texttt{M.reshape(...)} frequently returns a \textbf{view} of the original memory buffer rather than a copy. If you reshape a tensor to a matrix, perform an SVD, and then modify the matrix in-place, you risk silently corrupting the MPS state. To be safe, assign newly formed tensors directly rather than modifying slices:
	
	\begin{verbatim}
		# Safe: Assigns the new array object to the list index
		MPS[i] = U.reshape((D_left, d, D_new), order='C') 
		
		# Dangerous: Modifies the underlying memory buffer in-place
		MPS[i][:] = U.reshape((D_left, d, D_new), order='C') 
	\end{verbatim}
	
	\subsection{Robust SVD via SciPy}
	Always prefer \texttt{scipy.linalg.svd} over \texttt{numpy.linalg.svd}. SciPy defaults to the LAPACK divide-and-conquer driver (\texttt{gesdd}), which is substantially faster for the mid-sized matrices ($D \sim 1000$) encountered in DMRG. However, \texttt{gesdd} can occasionally fail to converge for highly degenerate matrices. A robust implementation catches this and falls back to the standard, highly stable \texttt{gesvd} driver.
	
	\begin{verbatim}
		from scipy.linalg import svd, LinAlgError
		
		try:
		U, S, Vh = svd(M, full_matrices=False)
		except LinAlgError:
		# Fallback to standard driver if divide-and-conquer fails
		U, S, Vh = svd(M, full_matrices=False, lapack_driver='gesvd')
	\end{verbatim}
	
	\subsection{The \texttt{np.einsum} Overhead in Iterative Solvers}
	While \texttt{np.einsum(..., optimize=True)} is highly readable for tensor contractions, it carries significant Python interpreter overhead because it calculates the optimal contraction path on every single call. Inside the \texttt{ApplyEffectiveHamiltonian} wrapper—which the Lanczos solver calls hundreds of times per site—you must replace \texttt{einsum} with explicit \texttt{np.tensordot} and \texttt{transpose} operations. This bypasses the Python interpreter and pushes the $\Order(D^3)$ heavy lifting directly to pre-compiled C/Fortran BLAS routines, frequently yielding a $>100\times$ speedup for the overall DMRG sweep.
	
	\subsection{Eigensolver Fallbacks for Exact Degeneracies}
	For small subsystem dimensions ($D \le 64$ at the boundaries), it is standard to bypass Lanczos and use dense exact diagonalization (\texttt{scipy.linalg.eigh}). However, for topologically ordered states like the AKLT model, exact ground-state degeneracies (e.g., from fractionalized edge spins) can cause the LAPACK divide-and-conquer driver to crash with \texttt{LinAlgError: Eigenvalues did not converge}. Robust codes must catch this exception and fall back to the general non-Hermitian solver (\texttt{scipy.linalg.eig}), sorting the real parts of the eigenvalues manually.
	
	\section{Implementation Verification Checklist}
	\label{sec:verification}
	
	The following phased testing protocol validates correct implementation of each subsystem before proceeding to full DMRG. Implement in order; failures at early stages indicate index ordering or conjugation bugs that must be fixed before testing variational optimization.
	
	\begin{table}[htbp]
		\centering
		\caption{Phased Implementation Tests}
		\label{tab:verification}
		\begin{tabularx}{\textwidth}{@{}l>{\raggedright\arraybackslash}p{3.5cm}X@{}}
			\toprule
			\textbf{Phase} & \textbf{Test} & \textbf{Expected Result} \\
			\midrule
			1. \texttt{Canonicalization} & 
			QR on random MPS $\to$ \textsc{LeftCanonicalize} & 
			Check $\sum_\sigma A^{(i)\sigma\dagger} A^{(i)\sigma} = \mathbb{I}_{D_i}$ for $i<L$ (within $10^{-14}$). Last site is vector $(D_{L-1}d) \times 1$. \\
			\midrule
			2. \texttt{Truncation} & 
			SVD compression of AKLT state ($D_{\max}=2$) & 
			$\eta = 0$ (exact representation maintained). Bond dimension remains $D=2$. \\
			\midrule
			3. \texttt{Environments} & 
			Build $L^{(i)}, R^{(i)}$ for AKLT with identity MPO & 
			Environments proportional to identity on auxiliary space. Verify $L^{(i)} \propto \delta_{a_{i-1},a'_{i-1}}$. \\
			\midrule
			4. \texttt{SingleSite} & 
			AKLT model, $L=10$, $D=2$, 2 sweeps & 
			Ground state energy $E_0 = -(L-1) \cdot 2/3 = -6.0$ (machine precision, $\Delta E < 10^{-12}$) using the explicit MPO from Sec.~\ref{sec:mpo_aklt}. \\
			\midrule
			5. \texttt{TwoSite} & 
			Heisenberg $L=20$, $D: 4\to20$, $\eps=10^{-10}$ & 
			Energy extrapolates to $-8.772$ (Bethe ansatz). Bond dimension adapts dynamically. \\
			\midrule
			6. \texttt{Convergence} & 
			Run single-site DMRG on gapped system & 
			Variance $\sigma^2 < 10^{-6}$ (strictest test). Check $\sigma^2 \sim D^{-2}$ scaling. \\
			\bottomrule
		\end{tabularx}
	\end{table}
	
	
	\subsection{Common Failure Modes}
	
	If tests fail, check these typical implementation errors:
	
	\begin{itemize}
		\item \textbf{Phase 1 fails (non-orthogonal $A$)}: Reshape index ordering incorrect. Verify row-major vs column-major convention (Section 1).
		\item \textbf{Phase 2 fails ($\eta > 0$ for AKLT)}: Incorrect SVD truncation logic or singular value sorting (descending order required).
		\item \textbf{Phase 3 fails (wrong environments)}: Missing complex conjugate on $A^{(i)*}$ or $B^{(i)*}$ in equations. Check environment indexing (off-by-one error $W^{(i)}$ vs $W^{(i\pm1)}$).
		\item \textbf{Phase 4 fails (wrong AKLT energy)}: MPO boundary conditions incorrect. Verify $b_{\text{left}}$ and $b_{\text{right}}$ select scalar component.
		\item \textbf{Phase 5 fails (stuck in local minimum)}: Insufficient bond dimension growth or missing noise injection.
		\item \textbf{Phase 6 fails (variance $\sigma^2$ large)}: Lanczos solver tolerance too loose or $\mathcal{H}_{\text{eff}}$ matrix-vector product bug (check einsum indices).
	\end{itemize}
	
	\subsection{Reference Values for Heisenberg Chain}
	For the $S=1/2$ Heisenberg model with open boundary conditions, exact diagonalization values for small $L$:
	\begin{center}
				\begin{tabular}{@{}cc@{}}
			\toprule
			$L$ & $E_0 / J$ \\
			\midrule
			4 & $-1.616025403784...$ \\
			6 & $-2.493577131515...$ \\
			8 & $-3.374932109083...$ \\
			10 & $-4.258035207...$ \\
			\bottomrule
		\end{tabular}
	\end{center}
	Use these to verify MPO construction before attempting DMRG on larger systems.
	
\end{document}